A Reproducible Data Analysis Workflow with RMarkdown, Git, Make & Docker

### 2020-05-21 manuscript.Rmd
added backticks
fix typo ì -> i

A `Makefile` obeys a simple scheme. It contains a list of recipes. Each recipe has a target (the name of the recipe) followed by a colon and a list of dependent targets or files. If any of the dependencies have changed since the last time the target was built, the recipe's commands are executed to recreate the target file. We illustrate the use of Makefiles with an example. Assume the final product is a manuscript (`manuscript.pdf`). This manuscript is written in Rmarkdown (`manuscript.Rmd`) and includes dynamically generated plots from a raw data file (`data/iris.csv`) that needs to be preprocessed first using a separate script (`R/prepare_data.R`) into a prepared data file (`ìris_prepped.csv`). A `Makefile` for these dependencies may look like this:

@@

data/ìris_prepped.csv: R/prepare_data.R data/iris.csv
  Rscript -e 'source("R/prepare_data.R")'
```

The first line after the comment is the first (default) target called 'all', which depends on manuscript.pdf, which itself is a target. If Make is called without an argument, the first target is built. To create `manuscript.pdf` (the second target in the file), the file `manuscript.Rmd` needs to be rendered, which depends on `data/ìris_prepped.csv`. This dependency is itself a target (the third target in the file). To create `data/ìris_prepped.csv`, `R/prepare_data.R` and `data/iris.csv` are needed. If you type `make manuscript.pdf`, Make first checks whether the dependencies do exist and, if not, creates them. Here, if `data/ìris_prepped.csv` does not exist, Make creates it by executing the third target. Also, if one of the dependencies of a target is newer than the target itself, Make updates everything that directly or indirectly depends on the target. Here, if `data\iris.csv` is newer than `data\ìris_prepped.csv`, Make will attempt to recreate `data\ìris_prepped.csv` first and then `manuscript.pdf`. If there is a dependency missing, and there is no target to make it, Make stops with an error message. It is a convention to have the first target named `all`, which creates the entire project. Subsequently, the command `make` without any argument automatically creates everything possible in the project. The button `Build All` from within RStudio triggers this process (see Figure \@ref(fig:buildpane-rstudio)).

@@

Finally, we conclude with some notes on project organization, which we think makes migrating projects to a reproducible workflow easier. The first step towards reproducibility is to create an R script or R Markdown file as the primary entry point for the analysis that runs on a local computer without error and performs the main statistical analyses. Next, one needs to make sure that all files relevant to the analysis can be moved to another computer. To this end, it is recommended that all files reside within a one folder (or enclosed subfolders within it) and all paths are relative to that folder because absolute paths are specific to a given computer. A robust solution to the problem of making sure that file access does not break across computing platforms are [RStudio projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` package [@here2017] to manage file access. The here package solves two common issues with relative paths. First, it takes care of the fact that path separator characters vary across operating systems (typically, slash or backslash). Second, it solves the issue that anchor points of relative paths may differ depending on context. For example, knitr interprets paths relative to the dynamic document, whereas R has a current working directory that may change over the course of an R session. The `here` package provides consistent paths relative to the project directory. The following three examples refer to local files ranging from absolute paths with system-specific path separators (bad) to relative paths using the `here` package:

>>>>

A `Makefile` obeys a simple scheme. It contains a list of recipes. Each recipe has a target (the name of the recipe) followed by a colon and a list of dependent targets or files. If any of the dependencies have changed since the last time the target was built, the recipe's commands are executed to recreate the target file. We illustrate the use of Makefiles with an example. Assume the final product is a manuscript (`manuscript.pdf`). This manuscript is written in Rmarkdown (`manuscript.Rmd`) and includes dynamically generated plots from a raw data file (`data/iris.csv`) that needs to be preprocessed first using a separate script (`R/prepare_data.R`) into a prepared data file (`iris_prepped.csv`). A `Makefile` for these dependencies may look like this:

@@

data/iris_prepped.csv: R/prepare_data.R data/iris.csv
  Rscript -e 'source("R/prepare_data.R")'
```

The first line after the comment is the first (default) target called 'all', which depends on manuscript.pdf, which itself is a target. If Make is called without an argument, the first target is built. To create `manuscript.pdf` (the second target in the file), the file `manuscript.Rmd` needs to be rendered, which depends on `data/iris_prepped.csv`. This dependency is itself a target (the third target in the file). To create `data/iris_prepped.csv`, `R/prepare_data.R` and `data/iris.csv` are needed. If you type `make manuscript.pdf`, Make first checks whether the dependencies do exist and, if not, creates them. Here, if `data/iris_prepped.csv` does not exist, Make creates it by executing the third target. Also, if one of the dependencies of a target is newer than the target itself, Make updates everything that directly or indirectly depends on the target. Here, if `data\iris.csv` is newer than `data\ìris_prepped.csv`, Make will attempt to recreate `data\ìris_prepped.csv` first and then `manuscript.pdf`. If there is a dependency missing, and there is no target to make it, Make stops with an error message. It is a convention to have the first target named `all`, which creates the entire project. Subsequently, the command `make` without any argument automatically creates everything possible in the project. The button `Build All` from within RStudio triggers this process (see Figure \@ref(fig:buildpane-rstudio)).

@@

Finally, we conclude with some notes on project organization, which we think makes migrating projects to a reproducible workflow easier. The first step towards reproducibility is to create an R script or R Markdown file as the primary entry point for the analysis that runs on a local computer without error and performs the main statistical analyses. Next, one needs to make sure that all files relevant to the analysis can be moved to another computer. To this end, it is recommended that all files reside within a one folder (or enclosed subfolders within it) and all paths are relative to that folder because absolute paths are specific to a given computer. A robust solution to the problem of making sure that file access does not break across computing platforms are [RStudio projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` package [@here2017] to manage file access. The `here` package solves two common issues with relative paths. First, it takes care of the fact that path separator characters vary across operating systems (typically, slash or backslash). Second, it solves the issue that anchor points of relative paths may differ depending on context. For example, knitr interprets paths relative to the dynamic document, whereas R has a current working directory that may change over the course of an R session. The `here` package provides consistent paths relative to the project directory. The following three examples refer to local files ranging from absolute paths with system-specific path separators (bad) to relative paths using the `here` package:
